### 2023-01-10 text/Supplements.Rmd
edit figure legends

---
title: Figure supplements
author: "Kei Jokura, Nobuo Ueda, Martin Gühmann, Luis Alfonso Yañez-Guerra, Kyle C. A. Wedgwood, Piotr Słowiński, Gáspár Jékely"
date: '*Living Systems Institute, University of Exeter, Stocker road, Exeter, EX4
  4QD, UK*'
output:
  pdf_document:
    latex_engine: lualatex
    template: default
  html_document:
    theme: lumen
  word_document: default
geometry: left=0.8in,right=0.8in,top=0.8in,bottom=0.5in
mainfont: Arial
monofont: Times New Roman
documentclass: article
classoption: onecolumn
linestretch: 1.2
fontsize: 10pt
bibliography: references.bib
csl: elife.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

![**Figure 1---figure supplement 1.** Phylogenetic tree of NOS by maximum likelihood. Tree robustness was tested with 1000 replicates of ultrafast bootstrap with the aLRT-SH-like and aBayes methods.](../figures/Fig1_sup1.png){width="100%"}

![**Figure 1---figure supplement 2.** Expression analysis of the NOS gene (magenta) using in situ HCR using the larva at 3-day-old in dorsal view (A) and anterior view (B). (B) Insert: showing two NOS cells on each side.](../figures/Fig1_sup2.png){width="100%"}

![**Figure 3---figure supplement 1.** **(A)** Overlaid trajectories for WT (left, n=20) and NOS mutant (right, n=31) 2-day-old larvae. Grey shows the trajectories before UV stimulation and purple shows the trajectories after UV stimulation. **(B)** The temporal changes in the vertical position of the WT and mutant 2-days-old larvae before and after UV stimulation are shown. The starting points of each larval trajectory are set to 0. After UV stimulation is indicated by purple squares. **(C)** Vertical swimming in wild-type (WT) and mutant (NOSΔ11 and NOSΔ23) larvae at 2-day-old stimulated with UV (395 nm) light from side, blue (488 nm) light from top and UV (395 nm) light from top. The data are shown in 30 s bins. **(D, E)** The temporal changes in the distance traveled of the WT and mutant in 2-day (D) and 3-day-old (E) larvae before and after UV stimulation are shown.](../figures/Fig3_sup1.png){width="100%"}

![**Figure 4---figure supplement 1.** Phylogenetic tree of guanylate cyclase by maximum likelihood (ML). Guanylate cyclase-coupled receptor and soluble guanylate cyclases (sGC) as outgroups. Guanylate cyclases with NIT domains are found in most animal phyla except Porifera, Ctenophora, Urochordata and Chordata.](../figures/Fig4_sup1.png){width="100%"}

![**Figure 4---figure supplement 2.** **(A)** Co-expression analysis image of the NIT-GC1 (magenta) and MLD-pedal2 amide proneuropeptide gene (MLD: green). Anterior view of the larva at 2-day-old. **(B, C)** Expression analysis of the NIT-GC2 gene (magenta) using in situ HCR. Anterior (B) and posterior (C) views of the larva at 3-day-old. **(D, E)** Localisation analysis for NIT-GC1 (B) and NIT-GC2 (C) morphant using NIT-GC1 and NIT-GC2 antibodies. Green shows co-staining with acetylated α-tubulin antibody (acTub). Anterior view of the larva at 2-day-old. **(F)** Co-localisation analysis using NIT-GC1 (magenta) and NOS (green) antibodies. Anterior view of the larva at 2-day-old.](../figures/Fig4_sup2.png){width="100%"}