### 2023-05-06 Manuscript/Manuscript.tex
upload ms in Manuscript/

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  10pt,
  onecolumn]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{setspace}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
  \setmainfont[]{Arial}
  \setmonofont[]{Times New Roman}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[left=0.8in,right=0.8in,top=0.8in,bottom=0.5in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Nitric-oxide-feedback to ciliary photoreceptor cells orchestrates UV avoidance in zooplankton},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Nitric-oxide-feedback to ciliary photoreceptor cells orchestrates
UV avoidance in zooplankton}
\author{true}
\date{}

\begin{document}
\maketitle

\setstretch{1.2}
\hypertarget{abstract}{%
\subsection{Abstract}\label{abstract}}

Nitric oxide (NO) generated by nitric-oxide synthase (NOS) is a key
regulator of animal physiology. Here we uncover a role for NO in shaping
circuit dynamics to orchestrate light-avoidance behaviour. We studied
UV-avoidance mediated by brain ciliary photoreceptors (cPRCs) in larval
\emph{Platynereis dumerilii}, a marine annelid. We found NOS
specifically expressed in interneurons (INNOS) postsynaptic to cPRCs.
Stimulation of cPRCs by violet light leads to cPRC inhibition but
concomitant INNOS activation and NO production. NO feeds back to cPRCs
and triggers their delayed activation through an unconventional guanylyl
cyclase. This results in the activation of projection interneurons and
the inhibition of serotonergic ciliomotor neurons. In \emph{NOS}
mutants, NO feedback and projection-neuron activation do not occur and
avoidance behaviour is defective. By mathematical modelling, we could
capture the circuit dynamics in both wild-type and mutant larvae. Our
results reveal how NO-mediated neuroendocrine signalling gates a
synaptic circuit to initiate light-avoidance behaviour.

\hypertarget{introduction}{%
\subsection{Introduction}\label{introduction}}

The ability to perceive changes in the environment and adapt behaviour
accordingly is essential for the survival of all living organisms.
Avoidance from UV radiation, which damages DNA essential for survival,
is proposed to be one of the behaviours that existed before the abundant
development of Earth's oxygen to block UV radiation in the atmosphere
(Pittendrigh). In the ocean, almost all zooplankton have a pronounced
rhythmic behaviour called Diel vertical migration (DVM), which consists
of upwards towards the surface at dusk and descend back to deeper water
before dawn. Avoidance behaviour to escape harmful UV radiation during
the day is closely linked to the control of vertical migration in the
ocean. However, little is known about the molecular neural circuit
mechanisms in UV avoidance behaviour in many animals, including
zooplankton. In neural circuits in general light response, the
activation of various receptors by excitatory and inhibitory
neurotransmitters such as glutamate, GABA and dopamine is known in
various species. In addition to this, it has long been known that in
nitric oxide (NO), as a neuromodulatory molecule, it is also important
as a modulator of visual activity. NO is a small free-radical molecule
with diffusible and changeable characteristics, and thus works as one of
the neural signalling molecules. The function of NO in the neural
circuitry of vision has been well studied in vertebrates. In the
mammalian retina, NO is an essential signalling molecule in all retinal
cells. NO produced by nitric oxide synthase (NOS) in amacrine cells, one
of the retinal cells, activates \emph{soluble guanylate cyclase (sGC)}
expressed in cone bipolar cells, which produces cGMP and regulates
downstream signalling. We therefore hypothesise that NO-mediated neural
signalling may be involved in the photoreceptor circuits that control UV
avoidance behaviour. To test this hypothesis we used the annelid
\emph{Platynereis dumerilii} \textbf{(Figure 1A)}, which conserved
vertebrate-type ciliary photoreceptors (cPRCs) (Arendt et al., 2004).
The cPRC expresses \emph{c-opsin1}, which absorbs UV radiation, is
involved in melatonin production and regulates circadian swimming
behaviour (Tosches et al., 2014; Tsukamoto et al., 2017). Furthermore,
we previously reported abnormal UV avoidance behaviour in
\emph{c-opsin1} knockout larvae (Verasztó et al., 2018). The cPRCs
express the UV-absorbing c-opsin1 and mediate UV avoidance in the larvae
(Verasztó et al., 2018). In adult worms, UV light and c-opsin1 modulate
locomotion and maturation (Veedin Rajan et al., 2021). These suggest
that Platynereis larvae are a highly suitable model for elucidating the
neural circuit mechanisms of NO-induced UV avoidance behaviour, and the
accessibility of genome editing, behavioural assays, calcium imaging and
whole-body connectomics could be very useful for study at the molecular
level(Ozpolat et al., 2021).

\hypertarget{results}{%
\subsection{Results}\label{results}}

\textbf{Nitric oxide synthase is expressed in interneurons of the
UV-avoidance circuit}

We identified a single \emph{nitric oxide synthase} (\emph{NOS}) gene in
the \emph{Platynereis dumerilii} genome and transcriptome data.
Phylogenetic analysis of NOS proteins indicate that \emph{Platynereis}
NOS belongs to an orthology group of metazoan NOS sequences
\textbf{(Figure 1---figure supplement 1)}. To characterise the
expression pattern of \emph{NOS} we used in situ hybridization chain
reaction (HCR) and transient transgenesis. In two- and three-day-old
larvae, we detected \emph{NOS} expression in four cells (two of them
weakly expressing) in the apical organ region \textbf{(Figure 1D and
Figure 1---figure supplement 2)}. The same four cells were labelled with
a reporter construct driving palmitoylated-tdTomato by a 12 kbp upstream
genomic region of the \emph{NOS} gene \textbf{(Figure 1E)}. This
reporter also revealed the axonal projections of the NOS-expressing
neurons. The position and morphology of the four \emph{NOS}-expressing
cells allowed us to identify the same cells in the three-day-old
whole-body volume EM data (Verasztó et al., 2020; Williams et al., 2017)
as four interneurons (INNOS) \textbf{(Figure 1B,C)}. The INNOS cells are
postsynaptic to the UV-sensory cPRCs and presynaptic to the INRGW
interneurons, which are also cPRC targets \textbf{(Figure 1C,F)}. INRGW
neurons synapse on the head serotonergic ciliomotor neurons (Ser-h1),
which synapse on the prototroch ciliary band and the cholinergic MC
ciliomotor neurons \textbf{(Figure 1C,F)}. We also detected \emph{NOS}
expression by HCR in the region of the visual eyes (adult eyes) and the
pigmented eyespots \textbf{(Figure 1---figure supplement 1)}.

\begin{figure}
\includegraphics[width=25in]{figures/Fig1} \caption{**Figure 1. Identification of NOS-expressing interneurons (INNOS) within the cPRC circuit.** **(A)** SEM image of a three-day-old Platynereis larva. **(B, C)** Five types of neurons (cPRC, INNOS, INRGW, Ser-h1 and MC) in the cPRC circuit reconstructed in the whole-body transmission electron microscopy (ssTEM) dataset of a three-day-old larva. Ciliated cells labelled with (B) are shown in grey. The INNOS on the ventral side labelled with (C) are shown in purple. Axons and dendrites appear as lines and cell body positions are represented by spheres. **(D)** Expression analysis of the NOS gene (magenta) using in situ HCR. Anterior view of the larva at two-day-old. Co-stained image using acetylated α-tubulin antibody (acTub: green). **(E)** Immunostaining with anti-HA antibody against HA tagged palmitoylated-tdTomato expressed under the upstream of the start site of NOS (NOSp::palmi-3xHA: magenta). Co-staining with acetylated α-tubulin antibody as marker for cilia of cPRC and axonal scaffold (acTub: green). Views of larvae at three-day-old from the anterior side. Insert: One of the reconstructed INNOS is shown from the anterior side. **(F)** Wiring diagram of the cPRC circuit. Hexagons and arrows represent cell groups and their synaptic connections, respectively. Numbers inside hexagons indicate the number of cells grouped in each node. Arrow line thickness (synapse weight) is equal to the number of synapses.}\label{fig:unnamed-chunk-2}
\end{figure}

\textbf{Nitric oxide produced during UV/violet stimulation gates the
output of the cPRC circuit}

The specific expression of \emph{NOS} in interneurons directly
postsynaptic to the cPRCs suggests that NO signalling is involved in the
UV response. To test if NO is produced during UV/violet responses, we
injected zygotes with the fluorescent NO-reporter DAF-FM and imaged
two-day-old larvae during the local stimulation of the cPRC cilia by 405
nm light. Following light stimulation, we detected an increase in DAF-FM
fluorescence in the region of the INNOS cells \textbf{(Figure 2)}.

\begin{figure}
\centering
\includegraphics[width=1\textwidth,height=\textheight]{../figures/Fig2.png}
\caption{\textbf{Figure 2. NO produced by UV/violet stimulation to
cPRCs.} \textbf{(A)} A NO-positive cell visualised by DAF-FM (green) is
shown (dashline). The white line indicates the outer frame of the larva.
All subsequent images are of two-day-old larvae. \textbf{(B)} Intensity
changes before (upper) and during 405 nm light stimulation (lower) in a
cell visualised by DAF-FM. \textbf{(C)} The intensity of DAF-FM changes
over time when cPRC (green) and other cells (ctr: gray) are stimulated
with UV/violet (405 nm) light. Purple squares indicate the timing of 405
nm light stimulation, the transparent lines show the optical responses
to individual intensities, each intensities are normalized (ΔF/F0) and
the dark trace lines and the shaded bands show the mean value and the
standard error range of the mean.}
\end{figure}

\textbf{Nitric oxide signalling mediates UV-avoidance behaviour}

To test this, we generated two \emph{Platynereis} \emph{NOS} knockout
lines with the CRISPR/Cas9 method. We recovered two deletions
(NOSΔ11/Δ11 and Δ23/Δ23), both leading to an early stop codon and thus
likely representing null alleles \textbf{(Figure 3A)}. We could
establish a homozygous line for both mutations indicating that NOS is
not an essential gene. To quantify UV avoidance, we recorded the
trajectories of freely swimming larvae in vertical columns, illuminated
laterally from two opposite sides with 395 nm UV/violet light. To
analyse the behaviour of larvae before and after UV stimulation, the
temporal changes in some parameters were compared using tracking of
individual larvae of the WT and mutant \textbf{(Figure 3B,C and Figure
3---figure supplement 1A-D)}. The results showed that in the WT there
was a rapid and strong increase in downward swimming approximately 15 s
after UV stimulation, whereas this did not occur in the mutant
\textbf{(Figure 3C and Figure 3---figure supplement 1B)}. Interestingly,
there was already a difference in swimming speed between the WT and the
mutant before UV stimulation (in the dark) \textbf{(Figure 3---figure
supplement 1C,D)}. After UV stimulation, the swimming speed of the WT
increased and reached a maximum after about 20 s. In contrast, the
mutants showed a slight increase in speed after UV stimulation, but
swimming speed was not maintained and returned to the speed before UV
stimulation. We also tested phototaxis, by exposing larvae to 480 nm
directional and collimated light from the top of the column. As
previously shown, wild-type larvae swim downward within 30 sec following
non-directional UV/violet light stimulation (Verasztó et al., 2018). In
contrast, homozygous \emph{NOS}-mutant larvae showed a strongly
diminished UV-avoidance response \textbf{(Figure 3D and Figure
3---figure supplement 1E)}. This phenotype is similar to the defective
UV-avoidance of \emph{c-opsin1} mutant larvae and reveals a requirement
for NOS in UV-avoidance behaviour (Verasztó et al., 2018).
\emph{NOS}-mutant larvae also showed reduced phototactic behaviour,
suggesting a function in the visual eyes mediating phototaxis (Randel et
al., 2014), where \emph{NOS} is also expressed \textbf{(Figure
1---figure supplement 1A,B)}. To test if the observed defects are due to
a role of NOS in neuronal signalling or circuit developmental, we also
tested larvae exposed to a NOS inhibitor, L-NAME. Larvae incubated for 5
min in 0.1 mM or 1 mM L-NAME showed defective UV avoidance, but normal
phototaxis \textbf{(Figure 3E)}. Overall, our results indicate an acute
requirement for NOS signalling in UV-avoidance and a possible indirect,
developmental role in the visual system. To investigate the role of NO
on UV avoidance behaviour, receptors for NO were explored.

\begin{figure}
\centering
\includegraphics[width=1\textwidth,height=\textheight]{../figures/Fig3.png}
\caption{\textbf{Figure 3. Larvae with inhibited NOS have a delayed
response to UV-avoidance.} \textbf{(A)} Top: The domain and Exon/Intron
structure of NOS. Bottom: Close-up region showing the genomic locus of
NOS gene and the wild-type sequence (WT) targeted by CRISPR/Cas9. The
generated mutants (NOSΔ11, NOSΔ23) are also shown. Pink indicates target
sites. Gray shows PAM sequences, red shows stop codons. \textbf{(B)}
Overlaid trajectories for WT (n=32) and NOS mutant (NOSΔ11/Δ11, n=26 and
NOSΔ23/Δ23, n=47) at three-day-old larvae. 0 sec as the starting point.
After 10 sec, UV (395 nm) stimulation from the side. \textbf{(C)} The
temporal changes in the vertical position of the WT and mutant larvae
before (dark) and after UV stimulation (side 395 nm) are shown. The
starting points of each larval trajectory are set to 0. \textbf{(D)}
Vertical swimming in wild-type (WT) and mutant (NOSΔ11 and NOSΔ23)
larvae at three-day-old stimulated with UV (395 nm) light from side,
blue (488 nm) light from top and UV (395 nm) light from top. The data
are shown in 30 s bins. \textbf{(E)} Vertical swimming in larvae treated
with NOS inhibitors, L-NAME at three-day-old stimulated with UV (395 nm)
light from side, blue (488 nm) light from top and UV (395 nm) light from
top. The data are shown in 30 s bins.}
\end{figure}

\textbf{Two unconventional guanylyl cyclases are expressed in the cPRCs}

NO generally acts via soluble guanylate cyclases (sGC), belonging to the
guanylate cyclase family with a CYC domain (PF00211). NO binding to the
heme group of sGC leads to increased cyclic guanosine monophosphate
(cGMP) production. Analysis of sGCs in \emph{Platynereis} indicated that
these genes are not expressed in any of the cells of the cPRC circuit
(Verasztó et al., 2017). Recently, Moroz and coworkers reported an
atypical but widely conserved family of guanylyl cyclases with a NIT
(nitrite/nitrate sensing) domain (PF08376) (NIT-GC) as potential
mediators of NO signalling (Moroz et al., 2020). To search for NIT-GCs
in \emph{Platynereis}, we searched transcriptome resources and retrieved
15 potential NIT-GC homologs \textbf{(Figure 4---figure supplement 1)}.
To analyse the relationship of these sequences to metazoan NIT-GCs, we
retrieved protein sequences with a CYC domain from the transcriptome and
genome databases of 45 metazoan species and 2 choanoflagellate species.
We carried out cluster analysis and did phylogenetic reconstruction on a
group of membrane-bound guanylyl cyclases with sGCs as an outgroup. In
agreement with Moroz et al. (Moroz et al., 2020),we found a group of GCs
with NIT domains with representatives in placozoans, cnidarians, some
ecdysozoans, echinoderms, and lophotrochozoans. The 15
\emph{Platynereis} sequences were parts of several major and ancient
clades of the tree \textbf{(Figure 4---figure supplement 1)}. To
characterise the expression of NIT-GCs, we using previously published
spatially mapped single-cell transcriptome data (Achim et al., 2015;
Williams et al., 2017). Among the 15 NIT-GCs, two showed high and
specific expression in the cPRCs and one was expressed in the INNOS
cells \textbf{(Figure 4---figure supplement 1)}. In the single-cell
data, we could identify the cPRCs by the specific expression of c-opsin1
and a pedal-peptide2 neuropeptide precursor (MLD proneuropeptide)
(Arendt et al., 2004; Williams et al., 2017), which have previously been
described as cPRC markers \textbf{(Figure 4---figure supplement 2A)}.
The INNOS cells were identified by NOS expression and their spatial
mapping. We decided to focus on two NIT-GCs expressed in the cPRCs and
with a full-length sequences, NIT-GC1 and NIT-GC2. To confirm the
single-cell data, we first carried out in situ HCR with probes for
\emph{NIT-GC1} and \emph{NIT-GC2} mRNA. Both genes showed very clear and
specific expression in the four cPRCs, as confirmed by co-labeling with
an acetylated α-tubulin antibody and with an HCR probe against \emph{MLD
proneuropeptide} \textbf{(Figure 4A,B and Figure 4---figure supplement
2A-C)}. To analyse the subcellular localisation of NIT-GC1 and NIT-GC2
at the protein level, we raised and affinity-purified polyclonal
antibodies against a specific peptide sequence from both proteins. In
immunostainings, we found that NIT-GC1 was localise to the region
corresponding to the axonal projections of the cPRCs in the anterior
nerve plexus \textbf{(Figure 4C)}. Co-immunostaining with the rabbit
NIT-GC1 and the rat NOS antibodies revealed that both proteins were
localised in close proximity in the neurosecretory plexus
\textbf{(Figure 4---figure supplement 2D)}. In contrast, NIT-GC2
strongly and specifically labelled the large sensory ciliary region of
the cPRCs \textbf{(Figure 4D)}. These different subcellular
localisations suggest that the two NIT-GCs are involved in different
intracellular signalling processes in the ciliary and axonal regions of
the cPRCs.

\textbf{NIT-GC1 is required for the depolarising response of cPRC to UV
stimulation}

To investigate how NIT-GCs-mediated NO signalling is involved in cPRC
activation, the calcium sensor GCaMP6s was ubiquitously expressed in
larvae and calcium imaging in the cPRC during UV/violet stimulation. As
we have shown previously, a 20-sec local stimulation of the ciliated
part of the cPRC with 405 nm light lead to a transient decrease in cPRC
calcium levels, likely due to the hyperpolarisation of the cPRCs
(Verasztó et al., 2018). After 20-sec, calcium levels in cPRCs were
raising again, reaching higher levels than at the start of the stimulus
- a response that may involve depolarisation. This activation phase
occurs after the end of the stimulation and may be due to a slower
neuroendocrine feedback (Verasztó et al., 2018). First, to determine
whether NO mediates such a feedback, we measured calcium levels in
\emph{NOS}-mutant larvae. While we detected the inhibition phase in cPRC
in \emph{NOS}-mutants, this was not followed by activation and calcium
levels remained low \textbf{(Figure 4E,F)}. These results revealed an
essential requirement for NO signalling in the second phase of the cPRC
response. Next, to characterise the role of NIT-GCs in changes of
calcium levels in cPRC, calcium imaging was performed when each
\emph{NIT-GC} gene was knocked down using morpholino \textbf{(Figure
4---figure supplement 2E,F)}. In NIT-GC1 morphant larvae, 405 nm light
stimulation caused a decrease in calcium levels in the cPRC, but the
increase in calcium levels that occurred after the light stimulation was
stopped was strongly suppressed \textbf{(Figure 4G)}. This result is
similar to the response of NOS mutants in cPRC, suggesting that NIT-GC1
mediates depolarising responses after UV stimulation of cPRC by NO and
may be involved in extracellular signalling. Interestingly, in NIT-GC2
morphant larvae, an increase in calcium levels occurred immediately
after light stimulation \textbf{(Figure 4H)}. The different responses of
cPRC upon knockdown of NIT-GC1 and NIT-GC2 suggest that there may be
different intracellular signalling mechanisms associated with their
respective localisations. We focused our analysis on NIT-GC1 because it
was localised to cPRC neuroaxons within the neuroplexas region and is
involved in extracellular signalling after UV stimulation. It is
generally known that depolarisation in the PRC involves elevated cGMPs
in the photoreceptors. We therefore used an in vitro system in which
cultured cells expressed NIT-GC1 to assess the cGMP activity of NIT-GC1.

\begin{figure}
\centering
\includegraphics[width=1\textwidth,height=\textheight]{../figures/Fig4.png}
\caption{\textbf{Figure 4. cPRC-specific expression of NIT-GC1 localises
in close proximity to NOS in the ANS region.} \textbf{(A, B)} Expression
analysis of the NIT-GC1 (C) and NIT-GC2 (D) gene (magenta) using in situ
HCR. Co-staining image using acetylated α-tubulin antibody (ac α-tub:
green). \textbf{(C, D)} Localisation analysis using NIT-GC1 and NIT-GC2
antibodies. Green shows co-staining with acetylated α-tubulin antibody
(ac α-tub). \textbf{(E, F)} Changes over time in GCaMP6s intensities in
cPRC when stimulated with UV in WT (dark yellow) and NOS mutant (E:
NOSΔ11/Δ11, F: NOSΔ23/Δ23) (light pink) larvae. \textbf{(G, H)} Changes
over time in GCaMP6s intensities in cPRC when stimulated with UV in
NIT-GC1 (G: NIT1) and NIT-GC2 (H: NIT2) knockdown by morpholino (MO1 and
2) larvae. \textbf{(I)} The domain structure of NIT-GC1. Below shows the
mutant NIT-GC1 (ΔNIT-GC1) with deletion of the NIT domain. Grey squares
indicate the transmembrane region (TM). \textbf{(J)} A schematic of the
NIT-GC1 assay; COS7 cells were used to express NIT-GC1 and the cGMP
indicator Green cGull, and the NO donor SNAP and DMSO as control. Green
cGull shows cGMP-dependent green fluorescence. \textbf{(K-N)} The
intensity changes in cGMP fluorescence over time for the four conditions
are shown. Gray traces show the optical responses to individual
intensities (n \textgreater{} 6). The individual intensities are
normalized (ΔF/F0). The dark line shows the mean value and the band the
standard error range of the mean. The respective solution was acted on 2
min after the start of imaging (grey bars).}
\end{figure}

\textbf{NIT-GC1 produces cGMP by nitric oxide}

NIT-GC1 has a highly conserved cyclase domain and activation of NIT-GC1
is expected to produce cGMP \textbf{(Figure 4I)}. Recently, a synthetic
protein called Green cGull was developed as a cGMP indicator that can be
used in cultured cell systems (Matsuda et al., 2016). We co-expressed
Green cGull and NIT-GC1 using the 2A self-cleaving peptide in cultured
cells COS-7 (monkey kidney) and tested the intracellular dynamics of
cGMP in the activation of NIT-GC1 \textbf{(Figure 4J)}. When the NO
donor SNAP was applied to the NIT-GC1 and Green cGull co-expression
cells, the intracellular cGMP levels was significantly increased
compared to when NIT-GC1 was not expressed \textbf{(Figure 4K,L)}. The
increase in fluorescence intensity also did not occur when DMSO was
applied \textbf{(Figure 4M)}. To confirm whether this increase in
intracellular cGMP levels is NIT domain-dependent response of NIT-GC1,
mutant form of NIT-GC1 without the NIT domain was expressed and SNAP was
applied \textbf{(Figure 4I)}. As a result, the increase in fluorescence
intensity was significantly reduced \textbf{(Figure 4N)}. These results
indicate that NIT-GC1 produces cGMP via the NIT domain. In bacteria, it
has been proposed that the NIT domain regulates cellular functions in
response to changes in extracellular and intracellular nitrate and
nitrite concentrations. Our results suggest that NIT-GC1 may be
sensitive to NO. This could be a possible feedback pathway in which
NIT-GC1 expressed in cPRC is activated by NO produced by INNOS. In order
to further explore the role of NO in the cPRC circuit, we therefore
compared calcium imaging of postsynaptic cells (INNOS, INRGW, Ser-h1 and
MC) in the cPRC circuit during UV stimulation of cPRC between WT and NOS
mutant larvae.

\textbf{Depolarisation of cPRC is required for activation of downstream
interneurons and ciliomotor neurons}

We next quantified the responses of the INNOS and INRGW following a
UV/violet stimulus to the cPRCs. To unambiguously identify the cells
from which we recorded calcium signals, we developed an on-slide
immunostaining method \textbf{(Figure 5---figure supplement 1)}. We used
the cell-specific neuropeptide markers using anti-RYamide (INNOS) and
anti-RGWamide (INRGW) antibodies (Conzelmann et al., 2011) to mark the
interneurons. Based on the position of the nuclei, we could correlate
live and fixed samples at a single-cell precision \textbf{(Figure
5A,B)}. In both wild-type and \emph{NOS}-mutant larvae, INNOS cells
showed increase calcium signals during the inhibition phase of the cPRCs
\textbf{(Figure 5D)}. In addition, INNOS depolarisation was suppressed
in NIT-GC2 molphants where cPRC hyperpolarisation was suppressed
\textbf{(Figure 5E)}. This indicates that cPRC hyperpolarisation is
required for INNOS depolarisation. In contrast, calcium signals
increased in INRGW cells in wild-type larvae during the second
activation phase of cPRCs, a response that was lacking in
\emph{NOS}-mutants \textbf{(Figure 5F,G)}. This suggests that there is
no circuit output to the Ser-h1 and MC cells and the ciliary band in
\emph{NOS}-mutants, consistent with the observed behavioural defects
\textbf{(Figure 5H,I)}.

\begin{figure}
\centering
\includegraphics[width=1\textwidth,height=\textheight]{../figures/Fig5.png}
\caption{\textbf{Figure 5. NO produced by UV stimulation to cPRCs
induces a depolarising response in cPRC via NIT-GC1.} \textbf{(A, B)}
Each left figure shows a calcium imaging using GCaMP6s. The respective
right figures show immunostaining images of the same positions in the
respective left figures, (A) shows RYamide antibodies, (B) shows
RGWamide and serotonin (5-HT) antibodies in red and DAPI in cyan.
Asterisks indicate the nuclei of the cPRC, respectively, and the nuclei
of the cells that were identified from the arrangement of the
surrounding cells are indicated by numbers. From the position of the
cells, the calcium imaging data identify INNOS in (A) and INRGW and
Ser-h1 in (B), respectively. \textbf{(C)} Correlation (Pearson's r) map
of neuronal activity of the INNOS, INRGW, Ser-h1 and MC neurons.
\textbf{(D)} Changes over time in GCaMP6s intensities in INNOS when cPRC
is stimulated with UV in WT (dark purple) and NOSΔ11/Δ11 mutant (light
purple) larvae. \textbf{(E)} Changes over time in GCaMP6s intensities in
INNOS when cPRC is stimulated with UV in NIT-GC2 morphants (MO1 and MO2:
pink and light pink) larvae. \textbf{(F)} Changes over time in GCaMP6s
intensities in INRGW when cPRC is stimulated with UV in WT (dark blue)
and NOSΔ23/Δ23 mutant (light blue) larvae. \textbf{(G)} Changes over
time in GCaMP6s intensities in INRGW when cPRC is stimulated with UV in
NIT-GC2 morphants (MO1 and MO2: blue and light blue). \textbf{(H, I)}
Changes over time in GCaMP6s intensities in Ser-h1 and MC when cPRC is
stimulated with UV in WT (Ser-h1 and MC: dark red and brown) and
NOSΔ11/Δ11 mutant (Ser-h1 and MC: light red and brown) larvae.}
\end{figure}

\hypertarget{discussion}{%
\subsection{Discussion}\label{discussion}}

In the present study, we explored the role of NO in the UV avoidance
behaviour of Platynereis. We demonstrated that there are interneurons
expressing NOS in the cPRC circuit and that NO is required for UV
avoidance in behavioural experiments using NOS mutants. We identified
the axon-localised type NIT-GC1 and the cilia-localised type NIT-GC2
expressed in the cPRCs as targets of NO and showed from cultured cell
experiments that these produce cGMP in response to NO. We have shown by
calcium imaging that INNOSs are activated in response to the
hyperpolarising response of the cPRCs, that the NO produced feeds back
to the cPRC via NIT-GC1 and induces a depolarising response in the cPRC,
and that the depolarising response of the cPRC is involved in activating
downstream INRGW. Lastly, we confirmed that RGWamide induces the
downward swimming behaviour of larvae.

\begin{figure}
\centering
\includegraphics[width=1\textwidth,height=\textheight]{../figures/Fig6.png}
\caption{\textbf{Figure 6. Gene expressions and signalling mechanisms
between cPRC, INNOS and INRGWa.} \textbf{(C)} Dot plot of genes
(columns) expressed in three types of cells (rows) in the cPRC circuit
using single cell RNA-Seq. The size of the dots is expressed in
proportion to the percentage of cells expressing that gene relative to
all cells. The colours represent the normal logarithm of the number of
transcripts in the cells expressing the gene. \textbf{(D)} Schematic
diagram of the signalling pathway of the cPRC circuit, focusing on the
NO feedback.}
\end{figure}

cPRCs expressing UV-absorbing c-opsin1 have high resting calcium levels
and show hyperpolarising and subsequent depolarising responses to UV
light (405 nm) (Verasztó et al., 2018). In the present study, we have
shown that the cells that strongly synapse with cPRCs are interneurons
that specifically express NOS (INNOS). The fluorescence intensity of
DAF-FM, which specifically detected INNOS, showed a transient tonic-like
increase dependent on UV stimulation of the cPRCs. The graph shows a
decrease in intensity after UV light is switched off, but the response
of DAF-FM to NO is irreversible and this decrease is due to breaching by
the relatively strong excitation light. Attempts to measure changes in
the intensity of the calcium indicator RGECO1a simultaneously with
DAF-FM imaging did not lead to detection of changes in fluorescence
intensity compared to GCaMP6s and could only visualise the cPRCs. In
calcium imaging using GCaMP6s, INNOS identified by the positional
relationship between the cPRC and INNOS and the labelling of RYa
depolarised in clear synchrony with the hyperpolarising response of the
cPRCs. This tonic response may involve dopaminergic nerves based on
single-cell data \textbf{(Figure 6)} {[}Puopolo et al. (2001){]}(Zhang
et al., 2007)(Randel et al., 2014). In the mammalian retina, it is known
that signals from bipolar photoreceptor cells are transmitted to
downstream NOS-expressing amacrine cells (NOAC), where NO is produced
(Jacoby et al., 2018). These suggest that UV stimulation of cPRCs causes
the production of NO via a depolarising response of INNOS.

Our results suggest that NIT-GC1, which is specifically expressed in
cPRCs, functions as a target for NO produced by INNOS and contributes to
the regulation of the cPRC circuit transmission pathway via a
depolarising response by increasing cGMP. In the mammalian retina, NO is
an essential signalling molecule and one of the most well-studied
(Cudeiro and Rivadulla, 1999). It has been reported that NO produced in
one of the retinal cells modulates cGMP-mediated neurotransmission by
activating presynaptic neural sGCs (general target of NO) through
retrograde signalling {[}Vielma et al. (2014){]}(Wei et al., 2012). In
Platynereis cPRC, NIT-GC1 was also identified as having the property of
producing cGMP in a NO-dependent manner. Furthermore, it is known that
CNGAα, one of the cyclic nucleotide-gated channels, is specifically
highly expressed in cPRC (Tosches et al., 2014), and in INNOS-cPRC
signalling, NO is implicated in the neuromodulation of the cPRC by
retrograde signalling. NO is a molecule with a very short half-life, and
when functioning as a signalling molecule, NOS and sGC are known to
localise very close together at synapses between neurons {[}Burette et
al. (2002){]}(Garthwaite, 2015). In our study, we found 12 sGCs in
Platynereis, but none expressed in the cPRC. Localisation analysis of
NOS and NIT-GC1 by antibody staining showed that the two proteins
localised very close together in the anterior neurosecretory plexus
region. Calcium imaging results using NOS mutant and NIT-GC1 knockdown
larvae suggest that NIT-GC1 activation induces a depolarising response
in the cPRC.

Another type of interneuron that strongly synapses with the cPRC is the
INRGWa, which synapses directly onto the ciliomotor neuron and is
therefore considered to be a very important neuron in the control of
larval behaviour {[}Verasztó et al. (2018){]}(Williams et al., 2017).
The present results suggest that the depolarising response of the cPRC
due to NO feedback induces INRGW activation. RGWamide is also a
neuropeptide that induces larval downward swimming quite strongly
(unpublished? need to show data?), and the delayed downward swimming of
NOS mutant larvae in response to UV stimulation may be responsible for
the failure of INRGW activation to occur. In addition, in many marine
invertebrates, NO has been reported to be involved in many aspects of
settlement behaviour and metamorphosis induction (Bishop and Brandhorst,
2007, 2003; Castellano et al., 2014; Leise et al., 2001; Pechenik et
al., 2007; Song et al., 2021; Ueda et al., 2016; Ueda and Degnan, 2014,
2013; Yang et al., 2018; Zhu et al., 2020). The downward swimming
behaviour of Platynereis larvae is the first step in the onset of
settlement behaviour and metamorphosis. These suggest that downward
swimming due to NO activation of INRGWs may trigger the onset of
settlement behaviour and metamorphosis {[}Conzelmann et al.
(2013b){]}(Williams et al., 2015). These suggest that NO feedback
signalling may act as a trigger for downward swimming via activation of
INRGWs, which induces the initiation of settlement behaviour and
metamorphosis.

An outstanding question is how the NO\textasciitilde NIT-GC response
alters neuronal dynamics. NO is a single molecule and, due to its short
half-life, cannot be transmitted over long distances and has a limited
signalling repertoire. It has therefore been considered unsuitable for
generating diversity in signalling compared to other signalling
molecules such as neuropeptides and classical neurotransmitters (Jékely,
2021). However, our results in the present study indicate that
NO-induced activation of NIT-GC1 is involved in the regulation of
neuronal circuits (cPRC pathway). The NIT domain was originally found in
bacteria and the existence of GCs with NIT domains was first reported by
Moroz et al in 2020 (Moroz et al., 2020; Shu et al., 2003). It has been
proposed that NIT domains in bacteria regulate cellular functions in
response to changes in extracellular and intracellular nitrate and/or
nitrite concentrations (Camargo et al., 2007). In fact, NO is a free
radical molecule, with a half-life of 5 msec in vivo, and is readily
converted to nitrate and nitrite (Garthwaite, 2015; Möller et al., 2019;
Santos et al., 2011). And nitrate and nitrite have been reported to
accumulate in cells and tissues with high NOS activity in placozoans and
cnidarians (Moroz et al., 2020, 2004). It is possible that the NIT
domains of NIT-GC1 identified in this study may also be used for
nitrate/nitrite sensing, as in bacteria. Furthermore, if they show
different sensitivities to NO, nitrite and nitrate, it is possible that
the temporal variation in their respective half-lives could give rise to
a repertoire of activation timings (Lundberg et al., 2011). In addition,
we have found at least 15 NIT-GCs in the Platynereis transcriptome data,
of which NIT-GC2 is highly expressed in the cPRC, in addition to
NIT-GC1. Interestingly, NIT-GC2 showed a very different localisation
pattern to NIT-GC1, and furthermore, the results of calcium imaging of
cPRC in response to UV stimulation using NIT-GC2 knockdown larvae were
the different dynamics (such that hyperpolarisation was suppressed) with
NIT-GC1 knockdown. It was not clear from the present analysis whether
this difference was due to differences in the localisation or the
property of the NIT-GCs. However, the diversification and spatially
different localisation patterns of NIT-GCs may increase their signalling
repertoire. We also believe that the appearance of NIT-GCs was one of
the means to generate the diversity of signalling pathways using NO
during the early stages of nervous system evolution in metazoans, since
sponges and ctenophores, where NO signalling is present, have no NIT-GCs
at all, as many as 12 have been found in placozoans, and NIT-GCs are
functional in the ``proto-eye'' circuit of Platynereis.

\hypertarget{materials-and-methods}{%
\subsection{Materials and Methods}\label{materials-and-methods}}

\textbf{CRISPR-Cas9 Design and Microinjection}

Before designing the small guide RNA (sgRNA) for the sgRNA:Cas9
nuclease, splice sites and polymorphic sites in our laboratory culture
were identified to avoid them. The sgRNA targeted the third exon of
Platynereis dumerilii NOS (target site: 5'-GGGCAATACTGGCTCCACTC-3'). The
sgRNA was assembled from two annealed oligonucleotides
(5'-TAGGGCAATACTGGCTCCACTC-3', 5'-AAACGAGTGGAGCCAGTATTGC-3') forming
overhangs for cloning into a BsaI site of the plasmid pDR27456 (Hwang et
al.~2013)(42250, Addgene), which contains next to the BsaI site a
tracrRNA sequence. The plasmid was then used to PCR amplify DNA
(primers: T7, 5'-AAAAGCACCGACTCGGTGCC-3') for synthesizing the sgRNA.
The DNA was purified with the QIAquick PCR Purification Kit (Qiagen).
From the DNA, the sgRNA was synthesized with the MEGAshortscript Kit
(Thermo Fisher Scientific) and was purified with the MEGAclear Kit
(Thermo Fisher Scientific). Cas9-mRNA was transcribed, capped, and
polyA-tailed with the mMessage mMachine Kit and the Poly(A) Tailing Kit
(both Thermo Fisher Scientific) from a plasmid (pUC57-T7-RPP2-Cas9)
containing the Cas9 ORF fused to 169 base pair 5' UTR from the
Platynereis dumerilii 60S acidic ribosomal protein P2. The sgRNA (18
ng/ml) and the Cas9-mRNA (180 ng/µl) were coinjected into fertilized
eggs of Platynereis dumerilii wild-type parents according to an
established injection procedure (Conzelmann et al., 2013a). The eggs
were kept at 18°C for 45 min before injection and were injected at
14.5°C. The injected individuals were kept at 18°C for 5 to 8 days in
6-well- plates (Nunc multidish no. 150239, Thermo Scientific) and then
cultured at 22°C until sexual maturity. The mature worms were crossed to
wild-type worms and the progeny was genotyped, resulting in two founder
lines, which were bred to homozygosity.

\textbf{NOS sequencing and genotyping}

For genotyping of the NOS locus, genomic DNA was isolated from single
larvae, groups of 6-20 larvae, or from the tails of adult worms. The DNA
was amplified by PCR (primers: 5'-GGTTCATTGGTTTCGATAACATTGCGG-3',
5'-CAGAGTCGATCAGTCTGCATATCTCCA-3') with the dilution protocol of the
Phusion Human Specimen Direct PCR Kit (Thermo Scientific). The PCR
product was sequenced directly with a nested sequencing primer
(5'-GGTGCTCTCCCGGGTACACAA-3'). A mixture of wild-type and deletion
alleles in a sample gave double peaks in the sequencing chromatograms,
with the relative height of the double peaks reflecting the relative
allele ratio in the sample.

\textbf{Vertical column setup for measuring photoresponses}

Photoresponses of larvae of different ages were assayed in a vertical
Plexiglas column (31 mm x 10 mm x 160 mm water height). The column was
illuminated from top with light from a monochromator (Polychrome II,
Till Photonics). The monochromator was controlled by AxioVision 4.8.2.0
(Carl Zeiss MicroImaging GmbH) via analog voltage. The light passed a
collimator lens (LAG-65.0-53.0-C with MgF2 Coating, CVI Melles Griot)
before entering the column. The column was illuminated from both sides
with light-emitting diodes (LEDs). The LEDs on each side were grouped
into two strips. One strip contained UV (395 nm) LEDs (SMB1W-395,
Roithner Lasertechnik) and the other infrared (810 nm) LEDs
(SMB1W-810NR-I, Roithner Lasertechnik). The UV LEDs were run at 4 V to
stimulate the larvae in the column from the side. The infrared LEDs were
run at 8 V (overvoltage) to illuminate the larvae for the camera (DMK
22BUC03, The Imaging Source), which recorded videos at 15 frames per
second and was controlled by IC Capture (The Imaging Source).

\textbf{Comparing behavior of wildtype and NOS-knockout 3-day-old
larvae}

To compare the behavior of wildtype and NOS-knockout larvae at 3 days in
the vertical column, the larvae were mixed and left in the dark for 5
min. The larvae were treated with NOS inhibitors for pharmacology. The
NOS inhibitors were L-NAME. The larvae were treated with different
concentrations in adjacent columns. The concentrations for the NOS
inhibitors were control, 1 mM, 0.1 mM. The larvae were recorded for 1
min in the dark followed by exposure to collimated cyan (480 nm) light
from the top of the column for 2 min, then 2 min darkness, and finally
collimated UV (395 nm) light from the top of the column for 2 min.
Stimulus light was provided by the monochromator (Polychrome II, Till
Photonics). Scripts are available at
\url{https://github.com/JekelyLab/NOS}.

\textbf{NOS Identification and Phylogenetic Analysis}

To identify NOS, we obtained a ``seed'' database of oxygenase domain in
Pfam database, PF02898. From these sequences, we produced a Hidden
Markov Model (HMM) and used this to mine the 47 metazoan species, 2
choanoflagellate species and 2 filasterea species investigated. HMM
models were run in HMMR3 with an e-value of 1e−15. We ran CD-Hit (Fu et
al., 2012) to eliminate redundant sequences (at a 80\% threshold). We
aligned the sequences with MAFFT version 7, with the iterative
refinement method E-INS-i. Alignments were trimmed with TrimAl in
gappy-out mode (Capella-Gutierrez et al., 2009). To calculate
maximum-likelihood trees, we used IQ-tree2 with the LG+G4 model. To
calculate branch support, we ran 1,000 replicates with the aLRT-SH-like
and aBayes methods (Minh et al., 2020). The sequences used for the
phylogenetic analysis are available in \textbf{Supplementary File 2},
the trimmed alignment is available in \textbf{Supplementary File 3}.

\textbf{NIT-GC Identification and Phylogenetic Analysis}

To identify NIT-GCs, we obtained a ``seed'' database of Adenylate and
Guanylate cyclase catalytic domain in Pfam database, PF00211. From these
sequences, we produced a Hidden Markov Model (HMM) and used this to mine
the 45 metazoan species, 2 choanoflagellate species and 2 filasterea
species investigated. HMM models were run in HMMR3 with an e-value of
1e−15. We ran CD-Hit (Fu et al., 2012) to eliminate redundant sequences
(at a 80\% threshold). The CLANS analysis is available as
\textbf{Supplementary File 1}. To identify clusters, we used the
convex-clustering option with 100 jack-knife replicates. The NIT-GCs are
extremely well conserved in membrane-bound guanylate cyclases and form
an easily recognizable cluster. To analyze the phylogeny of NIT-GCs, the
cluster containing these GCs together with membrane-bound guanylate
cyclases were parsed and used for tree building. We aligned the
sequences with MAFFT version 7, with the iterative refinement method
E-INS-i. Alignments were trimmed with TrimAl in gappy-out mode
(Capella-Gutierrez et al., 2009). To calculate maximum-likelihood trees,
we used IQ-tree2 with the LG+G4 model. To calculate branch support, we
ran 1,000 replicates with the aLRT-SH-like and aBayes methods (Minh et
al., 2020). The sequences used for the phylogenetic analysis are
available in \textbf{Supplementary File 2}, the trimmed alignment is
available in \textbf{Supplementary File 3}.

\textbf{Single-cell analysis}

We used Achim et al.~for the single-cell data (Achim et al., 2015). In
fact, in Williams et al., they used 107 cells as neurons by removing
duplicates from Achim et al.~single-cell data, so we used those cells
(Williams et al., 2017). Since the raw data were read count data, we
normalized them to TPM using Python. After that we converted them to
log10. From the sum of the expression levels in 107 cells for each gene,
I calculated the percentage in each cell. We chose the appropriate genes
for now. For each cell, we identified them with marker genes. After
created the data in Python, plotted it using R dot plots. RPKM
calculates the total number of reads per million bp and then divides by
the length of each gene, so it is not possible to compare between
samples. Instead, TPM first divides by the length of the gene and then
divides by the total number of reads per million bases, which allows for
more accurate comparisons between samples. In this case we wanted to
compare between samples, so we used TPM. In fact, TPM is becoming the
mainstream method for many single-cell analyses. The total TPM of each
gene between the samples was used to calculate the percentage of
expressed genes.

\textbf{In situ HCR}

Larvae were fixed and treated with Proteinase K, according to the
conventional WMISH protocol (Tessmar-Raible et al., 2005), with fixation
in 4\% paraformaldehyde/ PTW (PBS with 0.05\% Tween20) for 2 hr at room
temperature, and Proteinase K treatment in 100 µg/ml Proteinase K/ PTW
for 3 min (Tessmar-Raible et al., 2005). Specifically, for the HCR
protocol, samples were processed in 1.5 ml tubes. Probe hybridization
buffer, probe wash buffer, amplification buffer, and fluorescent HCR
hairpins were purchased from Molecular Instruments (Los Angeles, USA).
Hairpins associated with the b2 initiator sequence were labeled with
Alexa Fluor 647, and the hairpins associated with the b3 initiator
sequence were labeled with Alexa Fluor 546. To design probes for HCR, we
used custom software (Kuehn et al., 2021) to create 20 DNA oligo probe
pairs specific to P. dumerilii NOS, NIT-GC1, NIT-GC2, RYa-pNP (GenBank
accession: JF811330.1), c-opsin1 (GenBank accession: AY692353.1), CNGAα
(GenBank accession: KM199644.1), and MLD/pedal2-pNP (GenBank accession:
KF515945.1). The NOS, NIT-GC1 and NIT-GC2 probes were designed to be
associated with the b2 initiator sequence, while the RYa-pNP, c-opsin1,
CNGAα and MLD/pedal2-pNP probes were designed to be associated with the
b3 initiator sequence. For the detection stage, samples were
pre-hybridized in 200 µl of probe hybridization buffer for 1 hr at 37°C,
and then incubated in 250 µl hybridization buffer containing probe
oligos (4 pmol/ml) overnight at 37°C. To remove excess probe, samples
were washed 4× with 1 ml hybridization wash buffer for 15 min at 37°C,
and subsequently 2× in 1 ml 5× SSCT (5× SSC with 0.1\% Tween20) for 5
min at room temperature. For the amplification stage, samples were
pre-incubated with 100 µl of amplification buffer for 30 min, room
temperature, and then incubated with 150 µl amplification buffer
containing fluorescently labeled hairpins (40nM concentration (2ul of
3uM stock in 150ul amplification buffer, snap-cooled as described; (Choi
et al., 2018)) overnight in the dark at 25°C. To remove excess hairpins,
samples were washed in 1 ml 5× SSCT at room temperature, twice for 5
min, twice for 30 min, and once for 5 min. During the first 30 min wash,
samples were counterstained with DAPI (Cat. \#40043, Biotium, USA).

\textbf{Immunohistochemistry}

Whole-mount immunostaining of 2 day old Platynereis larvae fixed with
4\% paraformaldehyde were carried out using primary antibodies raised
against NIT-GC1, NIT-GC2, NOS, RYamide neuropeptide, RGWamide
neuropeptide in rabbit, plus a commercial antibody raised against
acetylated tubulin in mouse (Sigma T7451). The synthetic peptides
contained an N-terminal Cys that was used for coupling during
purification. Antibodies were affinity purified from sera as previously
described (Conzelmann and Jékely, 2012). Immunostainings were carried
out as previously described (Conzelmann and Jékely, 2012). The NOS
promoter (fragment sizes: 12 Kb) was amplified and cloned upstream of
3xHA- Palmi-tdTomato. Larvae injected with promoter constructs (ca. 250
ng/ml) were analysed for reporter expression at 3 days post
fertilization using an AxioImager Z.1 fluorescence wide-field microscope
(Carl Zeiss GmbH, Jena) and immediately fixed for immunostainings. The
protocol followed for immunostaining of HA-tagged reporters was recently
described (Verasztó et al., 2017). Specimens were imaged with a LSM 780
NLO or LSM 880 with Airysan Confocal Microscope (Zeiss, Jena).

\textbf{Calcium imaging}

For calcium imaging, 49--55 hpf larvae were used. Experiments were
performed at room temperature and larvae were immobilised by being
embedded in 2.5\% agarose filtered artificial seawater between a slide
and coverslip spaced with adhesive tape. GCaMP6s mRNA (1 mg/ml) was
injected into zygotes as described previously (Randel et al., 2014).
Larvae were imaged on a Zeiss LSM 880 with Airyscan (with a C-Apochromat
63X/1.2 Corr - water) with a frame rate of 1.88 frame/sec and an image
size of 512 x 512 pixels. The larvae were stimulated in a region of
interest (a circle with ?? pixel diameter) with 405 nm lasers controlled
by the Bleaching mode. The imaging laser had a similar intensity than
the stimulus laser but covered an area that was 10 times larger than the
stimulus ROI.

\textbf{Cell culture experiment}

Green cGull was used for the cGMP assay (Matsuda et al., 2016). A
full-length Pdum-NIT-GC1 and -NIT-GC2 coding sequences were amplified by
PCR starting from a \emph{Platynereis dumerilii} cDNA library and cloned
into the pcDNA3.1(+) vector using the T2A self-cleaving sequence. Cos-7
cells with low expression of endogenous soluble guanylate cyclase were
used as cultured cells for gene expression. This cell line was purchased
from Angio-proteomie (CAT no. cAP-0203). The Cos-7 cells were maintained
at 37 °C in 35mm dishes (Nunc™ Glass Bottom Dishes) containing 3 mL of
DMEM, high glucose glutamax medium (Thermo; Cat. No.~10566016)
supplemented with 10\% fetal bovine serum (Thermo; Cat. No.~10082147).
Upon reaching confluency of approximately 85\%, we transfected the cells
with the plasmid containing Green cGull-T2A-NITGC1. Transfections were
carried out with 150 ng of each plasmid and 0.3 μl of the transfection
Lipofectamine 3000 Reagent (invitrogen; Cat. No.~L3000001). Two days
post-transfection, we removed the culture medium and substituted it for
fresh DMEM-medium. For single-wavelength imaging experiments, cells in
35-mm dishes were washed twice and imaged in modified Ringer's buffer
(140 mM NaCl, 3.5 mM KCl, 0.5mM NaH2PO4 , 0.5mM S-3 MgSO4 , 1.5 mM CaCl2
, 10 mM HEPES, 2 mM NaHCO3 and 5 mM glucose). Dishes were mounted on a
stage heated at 37 °C and imaging was performed using an inverted
microscope (LSM880, Zeiss) equipped with an oil-immersion objective lens
(UApo/340, 40, NA = xx). Images were acquired using a xenon lamp,
460--495 nm excitation filter, 505-nm dichroic mirror and 510-- 550-nm
emission filter (????). S-Nitroso-N-acetyl-D, L-penicillamine (SNAP) was
purchased from Sigma-Aldrich (St.~Louis, MO, USA) . The exposure time of
the EM-CCD camera was controlled by the ZEN software (Zeiss). Images
were acquired every 15 s for 10 min and stimulation was initiated 2 min
after starting image acquisition. Imaging data analysis was performed
using ImageJ (National Institutes of Health, Bethesda, MD, USA).

\hypertarget{acknowledgements}{%
\subsection{Acknowledgements}\label{acknowledgements}}

This work was funded by the Wellcome Trust (214337/Z/18/Z).

\hypertarget{references}{%
\subsection*{References}\label{references}}
\addcontentsline{toc}{subsection}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-Achim2015}{}}%
Achim K, Pettit J-B, Saraiva LR, Gavriouchkina D, Larsson T, Arendt D,
Marioni JC. 2015. High-throughput spatial mapping of single-cell RNA-seq
data to tissue of origin. \emph{Nature Biotechnology}
\textbf{33}:503--509.
doi:\href{https://doi.org/10.1038/nbt.3209}{10.1038/nbt.3209}

\leavevmode\vadjust pre{\hypertarget{ref-Arendt2004}{}}%
Arendt D, Tessmar-Raible K, Snyman H, Dorresteijn AW, Wittbrodt J. 2004.
Ciliary Photoreceptors with a Vertebrate-Type Opsin in an Invertebrate
Brain. \emph{Science} \textbf{306}:869--871.
doi:\href{https://doi.org/10.1126/science.1099955}{10.1126/science.1099955}

\leavevmode\vadjust pre{\hypertarget{ref-bishop2007}{}}%
Bishop CD, Brandhorst BP. 2007. Development of nitric oxide
synthase-defined neurons in the sea urchin larval ciliary band and
evidence for a chemosensory function during metamorphosis.
\emph{Developmental Dynamics} \textbf{236}:1535--1546.
doi:\href{https://doi.org/10.1002/dvdy.21161}{10.1002/dvdy.21161}

\leavevmode\vadjust pre{\hypertarget{ref-bishop2003}{}}%
Bishop CD, Brandhorst BP. 2003. On nitric oxide signaling,
metamorphosis, and the evolution of biphasic life cycles.
\emph{Evolution and Development} \textbf{5}:542--550.
doi:\href{https://doi.org/10.1046/j.1525-142x.2003.03059.x}{10.1046/j.1525-142x.2003.03059.x}

\leavevmode\vadjust pre{\hypertarget{ref-burette2002}{}}%
Burette A, Zabel U, Weinberg RJ, Schmidt HHHW, Valtschanoff JG. 2002.
Synaptic Localization of Nitric Oxide Synthase and Soluble Guanylyl
Cyclase in the Hippocampus. \emph{The Journal of Neuroscience}
\textbf{22}:8961--8970.
doi:\href{https://doi.org/10.1523/jneurosci.22-20-08961.2002}{10.1523/jneurosci.22-20-08961.2002}

\leavevmode\vadjust pre{\hypertarget{ref-camargo2007}{}}%
Camargo A, Llamas A, Schnell RA, Higuera JoseJ́, Gonzaĺez-Ballester D,
Lefebvre PA, Fernańdez E, Galvań A. 2007. Nitrate Signaling by the
Regulatory Gene{\emph{NIT2}}in{\emph{Chlamydomonas}}. \emph{The Plant
Cell} \textbf{19}:3491--3503.
doi:\href{https://doi.org/10.1105/tpc.106.045922}{10.1105/tpc.106.045922}

\leavevmode\vadjust pre{\hypertarget{ref-Capella-Gutiuxe9rrez2009}{}}%
Capella-Gutierrez S, Silla-Martinez JM, Gabaldon T. 2009. trimAl: a tool
for automated alignment trimming in large-scale phylogenetic analyses.
\emph{Bioinformatics} \textbf{25}:1972--1973.
doi:\href{https://doi.org/10.1093/bioinformatics/btp348}{10.1093/bioinformatics/btp348}

\leavevmode\vadjust pre{\hypertarget{ref-castellano2014}{}}%
Castellano I, Ercolesi E, Palumbo A. 2014. Nitric Oxide Affects ERK
Signaling through Down-Regulation of MAP Kinase Phosphatase Levels
during Larval Development of the Ascidian Ciona intestinalis. \emph{PLoS
ONE} \textbf{9}:e102907.
doi:\href{https://doi.org/10.1371/journal.pone.0102907}{10.1371/journal.pone.0102907}

\leavevmode\vadjust pre{\hypertarget{ref-Choi2018}{}}%
Choi HMT, Schwarzkopf M, Fornace ME, Acharya A, Artavanis G, Stegmaier
J, Cunha A, Pierce NA. 2018. Third-generation {\emph{in situ}}
hybridization chain reaction: multiplexed, quantitative, sensitive,
versatile, robust. \emph{Development} \textbf{145}.
doi:\href{https://doi.org/10.1242/dev.165753}{10.1242/dev.165753}

\leavevmode\vadjust pre{\hypertarget{ref-Conzelmann2012}{}}%
Conzelmann M, Jékely G. 2012. Antibodies against conserved amidated
neuropeptide epitopes enrich the comparative neurobiology toolbox.
\emph{EvoDevo} \textbf{3}:23.
doi:\href{https://doi.org/10.1186/2041-9139-3-23}{10.1186/2041-9139-3-23}

\leavevmode\vadjust pre{\hypertarget{ref-Conzelmann2011}{}}%
Conzelmann M, Offenburger S-L, Asadulina A, Keller T, Münch TA, Jékely
G. 2011. Neuropeptides regulate swimming depth of {\emph{Platynereis}}
larvae. \emph{Proceedings of the National Academy of Sciences}
\textbf{108}.
doi:\href{https://doi.org/10.1073/pnas.1109085108}{10.1073/pnas.1109085108}

\leavevmode\vadjust pre{\hypertarget{ref-Conzelmann2013a}{}}%
Conzelmann M, Williams Elizabeth A, Krug K, Franz-Wachtel M, Macek B,
Jékely G. 2013a. The neuropeptide complement of the marine annelid
Platynereis dumerilii. \emph{BMC Genomics} \textbf{14}:906.
doi:\href{https://doi.org/10.1186/1471-2164-14-906}{10.1186/1471-2164-14-906}

\leavevmode\vadjust pre{\hypertarget{ref-Conzelmann2013}{}}%
Conzelmann M, Williams Elizabeth A, Tunaru S, Randel N, Shahidi R,
Asadulina A, Berger J, Offermanns S, Jékely G. 2013b. Conserved MIP
receptor{\textendash}ligand pair regulates {\emph{Platynereis}} larval
settlement. \emph{Proceedings of the National Academy of Sciences}
\textbf{110}:8224--8229.
doi:\href{https://doi.org/10.1073/pnas.1220285110}{10.1073/pnas.1220285110}

\leavevmode\vadjust pre{\hypertarget{ref-Cudeiro1999}{}}%
Cudeiro J, Rivadulla C. 1999. Sight and insight {\textendash} on the
physiological role of nitric oxide in the visual system. \emph{Trends in
Neurosciences} \textbf{22}:109--116.
doi:\href{https://doi.org/10.1016/s0166-2236(98)01299-5}{10.1016/s0166-2236(98)01299-5}

\leavevmode\vadjust pre{\hypertarget{ref-Fu2012}{}}%
Fu L, Niu B, Zhu Z, Wu S, Li W. 2012. CD-HIT: accelerated for clustering
the next-generation sequencing data. \emph{Bioinformatics}
\textbf{28}:3150--3152.
doi:\href{https://doi.org/10.1093/bioinformatics/bts565}{10.1093/bioinformatics/bts565}

\leavevmode\vadjust pre{\hypertarget{ref-garthwaite2015a}{}}%
Garthwaite J. 2015. From synaptically localized to volume transmission
by nitric oxide. \emph{The Journal of Physiology} \textbf{594}:9--18.
doi:\href{https://doi.org/10.1113/jp270297}{10.1113/jp270297}

\leavevmode\vadjust pre{\hypertarget{ref-jacoby2018}{}}%
Jacoby J, Nath A, Jessen ZF, Schwartz GW. 2018. A Self-Regulating Gap
Junction Network of Amacrine Cells Controls Nitric Oxide Release in the
Retina. \emph{Neuron} \textbf{100}:1149--1162.e5.
doi:\href{https://doi.org/10.1016/j.neuron.2018.09.047}{10.1016/j.neuron.2018.09.047}

\leavevmode\vadjust pre{\hypertarget{ref-Juxe9kely2021}{}}%
Jékely G. 2021. The chemical brain hypothesis for the origin of nervous
systems. \emph{Philosophical Transactions of the Royal Society B:
Biological Sciences} \textbf{376}:20190761.
doi:\href{https://doi.org/10.1098/rstb.2019.0761}{10.1098/rstb.2019.0761}

\leavevmode\vadjust pre{\hypertarget{ref-Kuehn2022}{}}%
Kuehn E, Clausen DS, Null RW, Metzger BM, Willis AD, Özpolat BD. 2021.
Segment number threshold determines juvenile onset of germline cluster
expansion in {\emph{Platynereis dumerilii}}. \emph{Journal of
Experimental Zoology Part B: Molecular and Developmental Evolution}
\textbf{338}:225--240.
doi:\href{https://doi.org/10.1002/jez.b.23100}{10.1002/jez.b.23100}

\leavevmode\vadjust pre{\hypertarget{ref-leise2001}{}}%
Leise EM, Thavaradhara K, Durham NR, Turner BE. 2001. Serotonin and
Nitric Oxide Regulate Metamorphosis in the Marine Snail{\emph{Ilyanassa
obsoleta}}. \emph{American Zoologist} \textbf{41}:258--267.
doi:\href{https://doi.org/10.1093/icb/41.2.258}{10.1093/icb/41.2.258}

\leavevmode\vadjust pre{\hypertarget{ref-lundberg2011}{}}%
Lundberg JO, Weitzberg E, Shiva S, Gladwin MT. 2011. The
nitrate{\textendash}nitrite{\textendash}nitric oxide pathway in mammals.
Humana Press. pp. 21--48.
doi:\href{https://doi.org/10.1007/978-1-60761-616-0_3}{10.1007/978-1-60761-616-0\_3}

\leavevmode\vadjust pre{\hypertarget{ref-Matsuda2017}{}}%
Matsuda S, Harada K, Ito M, Takizawa M, Wongso D, Tsuboi T, Kitaguchi T.
2016. Generation of a cGMP Indicator with an Expanded Dynamic Range by
Optimization of Amino Acid Linkers between a Fluorescent Protein and
PDE5? \emph{ACS Sensors} \textbf{2}:46--51.
doi:\href{https://doi.org/10.1021/acssensors.6b00582}{10.1021/acssensors.6b00582}

\leavevmode\vadjust pre{\hypertarget{ref-Minh2020}{}}%
Minh BQ, Schmidt HA, Chernomor O, Schrempf D, Woodhams MD, Haeseler A
von, Lanfear R. 2020. IQ-TREE 2: New Models and Efficient Methods for
Phylogenetic Inference in the Genomic Era. \emph{Molecular Biology and
Evolution} \textbf{37}:1530--1534.
doi:\href{https://doi.org/10.1093/molbev/msaa015}{10.1093/molbev/msaa015}

\leavevmode\vadjust pre{\hypertarget{ref-muxf6ller2019}{}}%
Möller MN, Rios N, Trujillo M, Radi R, Denicola A, Alvarez B. 2019.
Detection and quantification of nitric oxide{\textendash}derived
oxidants in biological systems. \emph{Journal of Biological Chemistry}
\textbf{294}:14776--14802.
doi:\href{https://doi.org/10.1074/jbc.rev119.006136}{10.1074/jbc.rev119.006136}

\leavevmode\vadjust pre{\hypertarget{ref-Moroz2004}{}}%
Moroz LL, Meech RW, Sweedler JV, Mackie GO. 2004. Nitric oxide regulates
swimming in the jellyfishAglantha digitale. \emph{The Journal of
Comparative Neurology} \textbf{471}:26--36.
doi:\href{https://doi.org/10.1002/cne.20023}{10.1002/cne.20023}

\leavevmode\vadjust pre{\hypertarget{ref-Moroz2020}{}}%
Moroz LL, Romanova DY, Nikitin MA, Sohn D, Kohn AB, Neveu E, Varoqueaux
F, Fasshauer D. 2020. The diversification and lineage-specific expansion
of nitric oxide signaling in Placozoa: insights in the evolution of
gaseous transmission. \emph{Scientific Reports} \textbf{10}.
doi:\href{https://doi.org/10.1038/s41598-020-69851-w}{10.1038/s41598-020-69851-w}

\leavevmode\vadjust pre{\hypertarget{ref-ozpolat2021}{}}%
Ozpolat BD, Randel N, Williams EA, Bezares-Calderón LA, Andreatta G,
Balavoine G, Bertucci PY, Ferrier DEK, Gambi MC, Gazave E,
Handberg-Thorsager M, Hardege J, Hird C, Hsieh Y-W, Hui J, Mutemi KN,
Schneider SQ, Simakov O, Vergara HM, Vervoort M, Jékely G,
Tessmar-Raible K, Raible F, Arendt D. 2021. The Nereid on the rise:
Platynereis as a model system. \emph{Zenodo}.
doi:\href{https://doi.org/10.5281/ZENODO.4907400}{10.5281/ZENODO.4907400}

\leavevmode\vadjust pre{\hypertarget{ref-pechenik2007}{}}%
Pechenik JA, Cochrane DE, Li W, West ET, Pires A, Leppo M. 2007. Nitric
Oxide Inhibits Metamorphosis in Larvae of{\emph{Crepidula fornicata}},
the Slippershell Snail. \emph{The Biological Bulletin}
\textbf{213}:160--171.
doi:\href{https://doi.org/10.2307/25066632}{10.2307/25066632}

\leavevmode\vadjust pre{\hypertarget{ref-puopolo2001}{}}%
Puopolo M, Hochstetler SE, Gustincich S, Wightman RM, Raviola E. 2001.
Extrasynaptic Release of Dopamine in a Retinal Neuron. \emph{Neuron}
\textbf{30}:211--225.
doi:\href{https://doi.org/10.1016/s0896-6273(01)00274-4}{10.1016/s0896-6273(01)00274-4}

\leavevmode\vadjust pre{\hypertarget{ref-Randel2014a}{}}%
Randel N, Asadulina A, Bezares-Calderón LA, Verasztó C, Williams EA,
Conzelmann M, Shahidi R, Jékely G. 2014. Neuronal connectome of a
sensory-motor circuit for visual navigation. \emph{eLife} \textbf{3}.
doi:\href{https://doi.org/10.7554/elife.02730}{10.7554/elife.02730}

\leavevmode\vadjust pre{\hypertarget{ref-santos2011}{}}%
Santos RM, Lourenço CF, Pomerleau F, Huettl P, Gerhardt GA, Laranjinha
J, Barbosa RM. 2011. Brain Nitric Oxide Inactivation Is Governed by the
Vasculature. \emph{Antioxidants \& Redox Signaling}
\textbf{14}:1011--1021.
doi:\href{https://doi.org/10.1089/ars.2010.3297}{10.1089/ars.2010.3297}

\leavevmode\vadjust pre{\hypertarget{ref-Shu2003}{}}%
Shu CJ, Ulrich LE, Zhulin IB. 2003. The NIT domain: a predicted
nitrate-responsive module in bacterial sensory receptors. \emph{Trends
in Biochemical Sciences} \textbf{28}:121--124.
doi:\href{https://doi.org/10.1016/s0968-0004(03)00032-x}{10.1016/s0968-0004(03)00032-x}

\leavevmode\vadjust pre{\hypertarget{ref-song2021}{}}%
Song H, Hewitt OH, Degnan SM. 2021. Arginine Biosynthesis by a Bacterial
Symbiont Enables Nitric Oxide Production and Facilitates Larval
Settlement in the Marine-Sponge Host. \emph{Current Biology}
\textbf{31}:433--437.e3.
doi:\href{https://doi.org/10.1016/j.cub.2020.10.051}{10.1016/j.cub.2020.10.051}

\leavevmode\vadjust pre{\hypertarget{ref-Tessmar-Raible2005}{}}%
Tessmar-Raible K, Steinmetz PRH, Snyman H, Hassel M, Arendt D. 2005.
Fluorescent two-color whole mount in situ hybridization in
{\emph{Platynereis dumerilii}} (Polychaeta, Annelida), an emerging
marine molecular model for evolution and development.
\emph{BioTechniques} \textbf{39}:460--464.
doi:\href{https://doi.org/10.2144/000112023}{10.2144/000112023}

\leavevmode\vadjust pre{\hypertarget{ref-Tosches2014}{}}%
Tosches M, Bucher D, Vopalensky P, Arendt D. 2014. Melatonin Signaling
Controls Circadian Swimming Behavior in Marine Zooplankton. \emph{Cell}
\textbf{159}:46--57.
doi:\href{https://doi.org/10.1016/j.cell.2014.07.042}{10.1016/j.cell.2014.07.042}

\leavevmode\vadjust pre{\hypertarget{ref-tsukamoto2017}{}}%
Tsukamoto H, Chen I-S, Kubo Y, Furutani Y. 2017. A ciliary opsin in the
brain of a marine annelid zooplankton is ultraviolet-sensitive, and the
sensitivity is tuned by a single amino acid residue. \emph{Journal of
Biological Chemistry} \textbf{292}:12971--12980.
doi:\href{https://doi.org/10.1074/jbc.m117.793539}{10.1074/jbc.m117.793539}

\leavevmode\vadjust pre{\hypertarget{ref-ueda2014}{}}%
Ueda N, Degnan SM. 2014. Nitric oxide is not a negative regulator of
metamorphic induction in the abalone haliotis asinina. \emph{Frontiers
in Marine Science} \textbf{1}.
doi:\href{https://doi.org/10.3389/fmars.2014.00021}{10.3389/fmars.2014.00021}

\leavevmode\vadjust pre{\hypertarget{ref-ueda2013}{}}%
Ueda N, Degnan SM. 2013. Nitric Oxide Acts as a Positive Regulator to
Induce Metamorphosis of the Ascidian Herdmania momus. \emph{PLoS ONE}
\textbf{8}:e72797.
doi:\href{https://doi.org/10.1371/journal.pone.0072797}{10.1371/journal.pone.0072797}

\leavevmode\vadjust pre{\hypertarget{ref-ueda2016}{}}%
Ueda N, Richards GS, Degnan BM, Kranz A, Adamska M, Croll RP, Degnan SM.
2016. An ancient role for nitric oxide in regulating the animal
pelagobenthic life cycle: evidence from a marine sponge.
\emph{Scientific Reports} \textbf{6}.
doi:\href{https://doi.org/10.1038/srep37546}{10.1038/srep37546}

\leavevmode\vadjust pre{\hypertarget{ref-Veedin2021}{}}%
Veedin Rajan VB, Häfker NS, Arboleda E, Poehn B, Gossenreiter T, Gerrard
E, Hofbauer M, Mühlestein C, Bileck A, Gerner C, Ribera d'Alcala M, Buia
MC, Hartl M, Lucas RJ, Tessmar-Raible K. 2021. Seasonal variation in UVA
light drives hormonal and behavioural changes in a marine annelid via a
ciliary opsin. \emph{Nature Ecology \& Evolution} \textbf{5}:204--218.
doi:\href{https://doi.org/10.1038/s41559-020-01356-1}{10.1038/s41559-020-01356-1}

\leavevmode\vadjust pre{\hypertarget{ref-Veraszto2018}{}}%
Verasztó C, Gühmann M, Jia H, Rajan VBV, Bezares-Calderón LA,
Piñeiro-Lopez C, Randel N, Shahidi R, Michiels NK, Yokoyama S,
Tessmar-Raible K, Jékely G. 2018. Ciliary and rhabdomeric
photoreceptor-cell circuits form a spectral depth gauge in marine
zooplankton. \emph{eLife} \textbf{7}.
doi:\href{https://doi.org/10.7554/elife.36440}{10.7554/elife.36440}

\leavevmode\vadjust pre{\hypertarget{ref-veraszto2020}{}}%
Verasztó C, Jasek S, Gühmann M, Shahidi R, Ueda N, Beard JD, Mendes S,
Heinz K, Bezares-Calderón LA, Williams E, Jékely G. 2020.
\href{http://dx.doi.org/10.1101/2020.08.21.260984}{Whole-animal
connectome and cell-type complement of the three-segmented
{\emph{Platynereis dumerilii}} larva}.

\leavevmode\vadjust pre{\hypertarget{ref-Veraszto2017}{}}%
Verasztó C, Ueda N, Bezares-Calderón LA, Panzera A, Williams EA, Shahidi
R, Jékely G. 2017. Ciliomotor circuitry underlying whole-body
coordination of ciliary activity in the Platynereis larva. \emph{eLife}
\textbf{6}.
doi:\href{https://doi.org/10.7554/elife.26000}{10.7554/elife.26000}

\leavevmode\vadjust pre{\hypertarget{ref-Vielma2014}{}}%
Vielma AH, Agurto A, Valdés J, Palacios AG, Schmachtenberg O. 2014.
Nitric Oxide Modulates the Temporal Properties of the Glutamate Response
in Type 4 OFF Bipolar Cells. \emph{PLoS ONE} \textbf{9}:e114330.
doi:\href{https://doi.org/10.1371/journal.pone.0114330}{10.1371/journal.pone.0114330}

\leavevmode\vadjust pre{\hypertarget{ref-Wei2012}{}}%
Wei T, Schubert T, Paquet-Durand F, Tanimoto N, Chang L, Koeppen K, Ott
T, Griesbeck O, Seeliger MW, Euler T, Wissinger B. 2012. Light-Driven
Calcium Signals in Mouse Cone Photoreceptors. \emph{Journal of
Neuroscience} \textbf{32}:6981--6994.
doi:\href{https://doi.org/10.1523/jneurosci.6432-11.2012}{10.1523/jneurosci.6432-11.2012}

\leavevmode\vadjust pre{\hypertarget{ref-Williams2015}{}}%
Williams EA, Conzelmann M, Jékely G. 2015. Myoinhibitory peptide
regulates feeding in the marine annelid Platynereis. \emph{Frontiers in
Zoology} \textbf{12}.
doi:\href{https://doi.org/10.1186/s12983-014-0093-6}{10.1186/s12983-014-0093-6}

\leavevmode\vadjust pre{\hypertarget{ref-williams2017}{}}%
Williams EA, Verasztó C, Jasek S, Conzelmann M, Shahidi R, Bauknecht P,
Mirabeau O, Jékely G. 2017. Synaptic and peptidergic connectome of a
neurosecretory center in the annelid brain. \emph{eLife} \textbf{6}.
doi:\href{https://doi.org/10.7554/elife.26349}{10.7554/elife.26349}

\leavevmode\vadjust pre{\hypertarget{ref-yang2018}{}}%
Yang X-X, Wong YH, Zhang Y, Zhang G, Qian P-Y. 2018. Exploring the
regulatory role of nitric oxide (NO) and the NO-p38MAPK/cGMP pathway in
larval settlement of the bryozoan {\emph{Bugula neritina}}.
\emph{Biofouling} \textbf{34}:545--556.
doi:\href{https://doi.org/10.1080/08927014.2018.1470240}{10.1080/08927014.2018.1470240}

\leavevmode\vadjust pre{\hypertarget{ref-zhang2007}{}}%
Zhang D-Q, Zhou T-R, McMahon DG. 2007. Functional Heterogeneity of
Retinal Dopaminergic Neurons Underlying Their Multiple Roles in Vision.
\emph{Journal of Neuroscience} \textbf{27}:692--699.
doi:\href{https://doi.org/10.1523/jneurosci.4478-06.2007}{10.1523/jneurosci.4478-06.2007}

\leavevmode\vadjust pre{\hypertarget{ref-zhu2020}{}}%
Zhu Y-T, Zhang Y, Liu Y-Z, Li Y-F, Yoshida A, Osatomi K, Yang J-L, Liang
X. 2020. Nitric oxide negatively regulates larval metamorphosis in
hard-shelled mussel (mytilus coruscus). \emph{Frontiers in Marine
Science} \textbf{7}.
doi:\href{https://doi.org/10.3389/fmars.2020.00356}{10.3389/fmars.2020.00356}

\end{CSLReferences}

\end{document}